---
vignette: >
  %\VignetteIndexEntry{Downsampling with group and stratum}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
litedown::reactor(
  print=NA,
  collapse = TRUE,
  comment = "#>",
  fig.width=10,
  fig.height=6)
data.table::setDTthreads(1)
```

In `mlr3`, a data set is represented as a `Task` with meta-data called roles which define how columns should be used for learning and evaluation.
The goal of this vignette is to explain how downsampling respects `group` and `stratum` roles.

# Introduction: role definitions

Previous column roles in `mlr3` are:

* `stratum`: proportions should be respected when splitting.
* `group`: rows that must stay together when splitting.

# Example using AZtrees data

The `AZtrees` data represent an image segmentation problem.
Each row is a pixel, which we want to classify as tree or not (`stratum` role).
Data were labeled by drawing polygons on a map (`group` role).
Data are divided into either 3 or 4 regions (`subset` role).

```{r}
data(AZtrees, package="mlr3resampling")
library(data.table)
options(datatable.print.keys=FALSE, datatable.print.class=FALSE)
AZtrees[, .(polygon, y)]
```

Above we see the two columns that are relevant for these roles in these data.
Below we count the number of rows for each polygon,

```{r}
dcast(AZtrees, polygon ~ y, length)[order(`Not tree`, -Tree)]
```

Above we see that there are several big polygons with no trees (hundreds of rows), whereas the polygons for presence of trees are smaller (max 45 rows).
Below we define a helper function which creates a `Task` and assigns roles based on arguments.

```{r}
trees <- function(..., folds=3L){
  trees_task <- mlr3::as_task_classif(AZtrees, target="y")
  role_list <- list(...)
  for(role in names(role_list)){
    trees_task$col_roles[[role]] <- role_list[[role]]
  }
  cv <- mlr3resampling::ResamplingSameOtherSizesCV$new()
  cv$param_set$values$folds <- folds
  cv$param_set$values$sizes <- 1
  cv$instantiate(trees_task)
  role_dt <- AZtrees[, .(y, polygon, set="ignored")]
  result_list <- list()
  for(it in 1:cv$iters){
    train_dt <- data.table(role_dt)[
      cv$train_set(it), set := "train"
    ][
      cv$test_set(it), set := "test"
    ][set!="test"]
    if("ignored" %in% train_dt$set){
      in_list <- list(
        rows=train_dt,
        groups=train_dt[, .(set=paste(unique(set), collapse=",")), by=.(y,polygon)])
      out_list <- list()
      for(unit in names(in_list)){
        in_dt <- in_list[[unit]]
        out_list[[unit]] <- rbind(in_dt[, table(y, set)], TOTAL=in_dt[, table(set)])
      }
      result_list[[length(result_list)+1]] <- out_list
    }
  }
  result_list
}
set.seed(1)
```

Above we set the seed to obtain reproducible results below (fold assignment is random).

## Cross-validation with no column roles

Below we show counts of rows and groups in each fold, when no column roles are set:

```{r}
trees()
```

We see in above that 

* in `rows`, there is about the same label distribution per fold (plus or minus 10 or 20 labels across folds), but not exactly the same (this can be achieved by setting `stratum` role). In `TOTAL` row we see the same number of rows per fold (plus or minus 1 because the total number of rows has a remainder when divided into the number of folds).
* in `groups`, rows from some polygons have been split into several different folds (`ignored,train` etc.), which is undesirable (and can be avoided by setting `group` role).

## Cross-validation using strata

Next, we illustrate the `stratum` column role.

```{r}
trees(stratum="y")
```

Above we see that

* in `rows`, both the label counts, and the `TOTAL`, are consistent across folds (plus or minus 1).
* in `groups`, there is large variation across folds, and some polygons have been split into different folds (column named `ignored,train` etc.).

## Cross-validation on polygons

Below we compute counts when setting `group` role to `polygon`:

```{r}
trees(group="polygon")
```

Above we see 

* in `rows`, there is substantial label distribution variation across folds (which can be avoided by setting `stratum` role). We see the `TOTAL` number of rows is equal across folds (again plus or minus 1).
* in `groups`, each polygon has been assigned to one fold (no `ignored,train` etc.), which is desirable. We see different numbers of `TOTAL` groups per fold, which is expected, because the algorithm attempts to equalize number or rows per fold, and there are different numbers or rows per group.

## Cross-validation on polygons with strata

Below we compute counts when setting `stratum` and `group` roles.

```{r}
trees(stratum="y", group="polygon")
```

Above we see

* in `rows`, the labels counts and `TOTAL` are consistent across folds (again plus or minus 1). TODO this is not true, need to fix code.
* in `groups`, each polygon has been assigned to one fold (no `1,2` etc.), which is desirable. We see different numbers of `TOTAL` groups per fold, which is expected, because the algorithm attempts to equalize number or rows per fold, and there are different numbers or rows per group.

# Example using respiratory data

These are medical data,

* with a binary target `outcome` (values 0 or 1),
* person `id` used for group.

TODO

# Conclusion

Overall these results show that the proposed resampling methods can handle complex cross-validation experiments, including downsampling with `group` and `stratum` roles.
