<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Benchmarking projects}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width=10,
  fig.height=6)
data.table::setDTthreads(1)
## output: rmarkdown::html_vignette above creates html where figures are limited to 700px wide.
## Above CSS from https://stackoverflow.com/questions/34906002/increase-width-of-entire-html-rmarkdown-output main-container is for html_document, body is for html_vignette
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this vignette is explain how to use `mlr3resampling::proj_*()` for computing machine learning benchmarks.

# Introduction: tasks, learners, resampling

In mlr3, a benchmark is defined as combinations of tasks, learners, and resampling iterations.

## Tasks

First we define a list of two tasks.

```{r}
task_list <- mlr3::tsks(c("spam", "german_credit"))
for(task_i in seq_along(task_list)){
  task <- task_list[[task_i]]
  task$col_roles$stratum <- task$col_roles$target
}
```

In the code above, we set `stratum` so that splits will preserve the target class proportions.

## Learners

Next, we define a list of two learners.

```{r}
learner_list <- list(
  mlr3::LearnerClassifFeatureless$new())
if(requireNamespace("rpart")){
  learner_list$rpart <- mlr3::LearnerClassifRpart$new()
}
for(learner_i in seq_along(learner_list)){
  L <- learner_list[[learner_i]]
  L$predict_type <- "prob"
}
```

In the code above, we use the `prob` predict type, because we want to use ROC-AUC as an evaluation metric:

```{r}
meas_list <- mlr3::msrs(c("classif.auc","classif.acc"))
```

## Resampling

Finally we create an instance of 3-fold CV, which we will use as the train/test splitting method.

```{r}
kfoldcv <- mlr3::rsmp("cv", folds=3)
```

# `mlr3::benchmark_grid()` and `mlr3resampling::proj_grid()`

The code in the previous section is required in both TODO

```{r}
bgrid <- mlr3::benchmark_grid(task_list, learner_list, kfoldcv)
```

```{r}
pdir <- tempfile()
pgrid <- mlr3resampling::proj_grid(
  pdir, task_list, learner_list, kfoldcv,
  score_args = meas_list)
```

# testing one job

```{r}
rdir <- tempfile()
batchtools::makeExperimentRegistry(rdir)
mlr3batchmark::batchmark(bgrid)
batchtools::testJob(1)
```

```{r}
mlr3resampling::proj_test(pdir, max_jobs=1)
```

# testing one job for each algo and data set

A simple way to test each algorithm and data set with batchtools is via the code below, which uses `repl==1` to consider only the first cross-validation iteration:

```{r}
jt <- batchtools::getJobTable()[repl==1]
sapply(jt$job.id, batchtools::testJob)
```

There is no error above, but it is not clear if the result is reasonable.
Another way to do it with batchtools is to submit a subset of jobs, as below:

```{r}
batchtools::submitJobs(jt$job.id)
batchtools::waitForJobs()
test_res <- mlr3batchmark::reduceResultsBatchmark(jt$job.id)
test_res$score(meas_list)
```

The result above includes test AUC and accuracy values, which are more convincing test.

Our proposed way to test is via:

```{r}
mlr3resampling::proj_test(pdir)
```

The results above include test AUC and accuracy, but with different values because the tasks are down-sampled to make training faster. For even faster runs, Learners may define `edit_learner()` method, which edits the learner to make training faster. For example, the default with `AutoTuner` and `TorchLearner` is to only use two epochs of training during `proj_test()`, so you can quickly see if results look reasonable, before running training using the full number of epochs, and the full data set.
The code below reads the down-sampled tasks:

```{r}
rds.vec <- Sys.glob(file.path(pdir,"test","tasks","*rds"))
for(rds.i in seq_along(rds.vec)){
  mini_task <- readRDS(rds.vec[[rds.i]])
  print(table(mini_task$data()[[1]]))
}
```

The output above shows that there are at least 10 data per class in the down-sampled tasks.

# running locally

```{r}
```
