<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Benchmarking projects}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width=10,
  fig.height=6)
data.table::setDTthreads(1)
options(width=120)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this vignette is to compare and contrast calculation of machine learning benchmarks using

* `mlr3` and `mlr3batchmark` packages, which are the built-in/previous/existing methods.
* `mlr3resampling::proj_*()` functions, which are the new methods proposed in this package.

# Introduction: tasks, learners, resampling

In mlr3, a benchmark is defined as combinations of tasks, learners, and resampling iterations.
The code in this section must be run using both methods (previous and proposed).

## Tasks

First we define a list of two tasks.

```{r}
task_list <- mlr3::tsks(c("spam", "german_credit"))
for(task_i in seq_along(task_list)){
  task <- task_list[[task_i]]
  task$col_roles$stratum <- task$col_roles$target
}
```

In the code above, we set `stratum` so that splits will preserve the target class proportions.

## Learners

Next, we define a list of two learners.

```{r}
learner_list <- list(
  mlr3::LearnerClassifFeatureless$new())
if(requireNamespace("rpart")){
  learner_list$rpart <- mlr3::LearnerClassifRpart$new()
}
for(learner_i in seq_along(learner_list)){
  L <- learner_list[[learner_i]]
  L$predict_type <- "prob"
}
```

In the code above, we use the `prob` predict type, because we want to use ROC-AUC as an evaluation metric, in addition to accuracy:

```{r}
measure_list <- mlr3::msrs(c("classif.auc","classif.acc"))
```

## Resampling

Finally we create an instance of 3-fold CV, which we will use as the train/test splitting method.

```{r}
kfoldcv <- mlr3::rsmp("cv", folds=3)
```

# Define the grid of combinations

In this section, we explain the two alternative methods for defining the grid of combinations.

## Previous method, `mlr3::benchmark_grid()`

In the code below, we must save the result, called `bgrid`, which exists only in R (not yet saved to the file system).

```{r}
(bgrid <- mlr3::benchmark_grid(task_list, learner_list, kfoldcv))
```

After that, we save the grid to the file system via:

```{r}
reg_dir <- tempfile()
reg <- batchtools::makeExperimentRegistry(reg_dir)
mlr3batchmark::batchmark(bgrid)
```

## Proposed method, `mlr3resampling::proj_grid()`

In the code below, the grid of combinations is saved to the `proj_dir` directory.

```{r}
proj_dir <- tempfile()
mlr3resampling::proj_grid(
  proj_dir, task_list, learner_list, kfoldcv,
  score_args = measure_list)
```

At this step in the proposed code above,

* The first argument `proj_dir` defines the project directory, in which data and result files are saved. This is the analog of the registry directory, `reg_dir`, in the previous approach.
* The named argument `score_args = measure_list` defines the evaluation metrics we will use to measure the accuracy of predictions on the held-out test sets. Each parallel job will compute these evaluation metrics for one train/test split and one Learner.

# Testing

After defining the grid of combinations, we typically want to do a small test on the local system, to make sure there are no errors, before submitting the full computation.

## Testing one job

To test one job using the previous approach, we use:

```{r}
batchtools::testJob(1)
```

To test one job using the proposed approach, we use:

```{r}
mlr3resampling::proj_test(proj_dir, max_jobs=1)
```

We tested a featureless learner, which is not a great test (other learners may fail).
So typically we run a more extensive test, as in the next section.

## Test one job for each algo and data set

A simple way to test each algorithm and data set with batchtools is via the code below, which uses `repl==1` to consider only the first cross-validation iteration, run on the local machine:

```{r}
jt <- batchtools::getJobTable()
jt1 <- jt[repl==1]
sapply(jt1$job.id, batchtools::testJob)
```

There is no error above, but it is not clear if the result is reasonable.
Another way to do it with batchtools is to submit a subset of jobs, as below:

```{r}
batchtools::submitJobs(jt1$job.id)
batchtools::waitForJobs()
test_res <- mlr3batchmark::reduceResultsBatchmark(jt1$job.id)
test_res$score(measure_list)
```

The result above includes test AUC and accuracy values, which are more convincing test (but this involves actually submitting jobs on the cluster).

Our proposed way to test is via:

```{r}
mlr3resampling::proj_test(proj_dir)
```

The results above include test AUC and accuracy, but with different values because the tasks are down-sampled to make training faster.

# Running all jobs locally (small benchmarks)

For some small benchmarks, you may want to compute all results on your local machine (not a cluster).
For both the previous and proposed approaches, we use the code below to enable computation in parallel using all the CPUs on the local machine.

```{r}
future::plan("multisession")
```

The previous approach uses the code below:

```{r}
bench_result <- mlr3::benchmark(bgrid)
bench_score <- bench_result$score(measure_list)
bench_score[, .(task_id, learner_id, iteration, classif.auc, classif.acc)]
```

The proposed approach uses the code below:

```{r}
proj_score <- mlr3resampling::proj_compute_all(proj_dir)
proj_score[, .(task_id, learner_id, iteration, classif.auc, classif.acc)]
```

The results are slightly different because the cross-validation is instantiated with different random seeds.

# Running all jobs on a cluster (large benchmarks)

In the previous system, we use batchtools.
Various parallel computation methods are supported.
Prior to running the code below on the cluster, you should create a `~/.batchtools.conf.R` file, as explained in [my R batchtools on Monsoon tutorial](https://tdhock.github.io/blog/2020/monsoon-batchtools/).

TODO conf in extdata
test if slurm prog present
set env var here to look there for config

```{r}
batchtools::submitJobs(jt$job.id)
batchtools::waitForJobs()
test_res <- mlr3batchmark::reduceResultsBatchmark(jt$job.id)
test_res$score(measure_list)
```

The computation above was in 

# Comparison

TODO , which will be saved to `results.csv` in the last job. Using the previous approach, `measure_list` is provided after the computations have run, so must be computed in a final combination step, after the last job.

Testing is more convenient

# New features

## Quick testing via Task down-sampling and Learner `edit_learner()` method

For even faster runs, Learners may define `edit_learner()` method, which edits the learner to make training faster. For example, the default with `AutoTuner` and `TorchLearner` is to only use two epochs of training during `proj_test()`, so you can quickly see if results look reasonable, before running training using the full number of epochs, and the full data set.

The code below reads the down-sampled tasks:

```{r}
rds.vec <- Sys.glob(file.path(proj_dir,"test","tasks","*rds"))
for(rds.i in seq_along(rds.vec)){
  mini_task <- readRDS(rds.vec[[rds.i]])
  print(table(mini_task$data()[[1]]))
}
```

The output above shows that there are at least 10 data per class in the down-sampled tasks.

## Model interpretation via `save_learner()` method
