---
vignette: >
  %\VignetteIndexEntry{Reproducible benchmarks}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
litedown::reactor(
  collapse = TRUE,
  comment = "#>",
  fig.width=10,
  fig.height=6)
data.table::setDTthreads(1)
```

The goal of this vignette is to explain how to compute reproducible machine learning benchmarks.

# Introduction

Reproducibility is the ability to re-compute the exact same results, given the exact same inputs, possibly on a variety of different computers.

In mlr3 benchmarks, there are three components, all of which may present barriers to reproducibility:

* Tasks represent data sets, which may not have pre-defined splits or cross-validation folds, in which case mlr3 will create them randomly (with a loss of reproducibility).
* Learners represent algorithms which may be deterministic (reproducible) or random (not reproducible unless random seed is set).
* Resampling methods like cross-validation are inherently random.

When using `mlr3resampling::proj_grid()`, the default is `train_seed=1L`, which means R’s random seed will be set before training.

For reproducible train/test splits, we recommend

* randomly creating a fold column, so that fold values respect strata such as class labels.
* save data with the fold column as a new CSV file, which can be used to communicate to others what cross-validation splits you used in your benchmark.
* assign the fold column to the `fold` role in the mlr3 Task.
* use `mlr3resampling::ResamplingSameOtherSizesCV` which uses the column with `fold` role in cross-validation.

These steps ensure that the benchmark results are reproducible, given the CSV file with fold column, and the random seed for training.

# Example

We begin by defining the Resampling method (default is 3-fold CV).

```{r}
kfold <- mlr3resampling::ResamplingSameOtherSizesCV$new()
```

Next we load the `spam` binary classification task, optionally down-sampling to 200 rows to make the computations quicker.

```{r}
spam <- mlr3::tsk("spam")
## uncomment next line to speedup rendering:
#spam$filter(as.integer(seq(1, spam$nrow, length.out = 200)))
spam
```

Next, we create a new CSV file with fold IDs.

```{r}
library(data.table)
spam_with_fold.csv <- if(interactive())"~/spam_with_fold.csv" else tempfile()
spam_with_fold.dt <- spam$data()[
, Fold := rep(1:3, length.out = .N)
, by = type]
fwrite(spam_with_fold.dt, spam_with_fold.csv)
spam_with_fold.dt[, table(Fold, type)]
```

The outpat above shows that the number of samples in each class is approximately constant across folds.
Next, we use these data to define a new task with the `fold` role.

```{r}
spam_with_fold <- mlr3::TaskClassif$new(
  "spam_with_fold", spam_with_fold.dt, target="type")
spam_with_fold$col_roles$fold <- "Fold"
spam_with_fold$col_roles$feature <- spam$col_roles$feature
```

Below we assign the `stratum` role to both tasks, for proportional down-sampling.

```{r}
spam_with_fold$col_roles$stratum <- c("type","Fold")
spam$col_roles$stratum <- "type"
```

Next we define learners and ensure their predict types are real-valued scores (`predict_type="prob"`), so we can compute AUC.

```{r}
learner_list <- list(
  mlr3learners::LearnerClassifCVGlmnet$new(),
  mlr3::LearnerClassifRpart$new())
for(learner.i in seq_along(learner_list)){
  learner_list[[learner.i]]$predict_type <- "prob"
}
```

Next, we create a project grid.

```{r}
pdir <- if(interactive())"~/pdir" else tempfile()
task_list <- list(spam, spam_with_fold)
unlink(pdir, recursive = TRUE)
measure_list <- mlr3::msrs("classif.auc")
mlr3resampling::proj_grid(
  pdir, task_list, learner_list, kfold, score_args=measure_list)
```

## Demonstration with test project

Below we run `proj_test()` twice.

```{r}
test_res_list <- list()
for(run.num in 1:2){
  tres <- mlr3resampling::proj_test(pdir, min_samples_per_stratum = 20)
  test_res_list[[run.num]] <- data.table(
    run=paste0("run", run.num), tres$results.csv)
}
test_res <- rbindlist(test_res_list)[
, algorithm := sub("classif.", "", learner_id)]
dcast(
  test_res, task_id + algorithm ~ run, value.var="classif.auc")
```

The result above shows that the two runs have the same AUC values only for `spam_with_fold`, not for `spam` (which gets a different fold vector for each run).

## Demonstration with full benchmark

Below we create two project grids from the same code.

```{r}
set.seed(1)
jobs_list <- list()
for(run.num in 1:2){
  pdir <- if(interactive())paste0("~/pdir",run.num) else tempfile()
  unlink(pdir, recursive = TRUE)
  pgrid <- mlr3resampling::proj_grid(
    pdir, task_list, learner_list, kfold, score_args=measure_list)
  jobs_list[[run.num]] <- data.table(
    run=paste0("run", run.num), pdir, job.i=pgrid[
    , which(test.fold==1 & grepl("glmnet", learner_id))])
}
```

In the code above, the two grids have different random fold assignments.
Below, we compute one glmnet job for each run and task.

```{r}
proj_res <- rbindlist(jobs_list)[
, mlr3resampling::proj_compute(job.i, pdir)
, by=.(run, pdir, job.i)][
, algorithm := sub("classif.", "", learner_id)]
dcast(
  proj_res, task_id + algorithm ~ run, value.var="classif.auc")
```

The results above are consistent with the results above.

# Comparison with batchtools

If you want consistent results between runs with batchtools, you need to set the seed argument when making the registry.

```{r}
(bgrid <- mlr3::benchmark_grid(task_list, learner_list, kfold))
if(requireNamespace("mlr3batchmark")){
  batch_dt_list <- list()
  for(run.num in 1:2){
    reg_dir <- if(interactive())paste0("~/reg",run.num) else tempfile()
    unlink(reg_dir, recursive = TRUE)
    reg <- batchtools::makeExperimentRegistry(reg_dir, seed=1)
    mlr3batchmark::batchmark(bgrid)
    jt <- batchtools::getJobTable()
    jt1 <- jt[repl==1]
    batchtools::submitJobs(jt1)
    batchtools::waitForJobs()
    ignore.learner <- function(L){
      L$learner_state$model <- NULL
      L
    }
    bt_res <- mlr3batchmark::reduceResultsBatchmark(jt1, fun=ignore.learner)
    bt_score <- bt_res$score(measure_list)
    batch_dt_list[[run.num]] <- data.table(
      run=paste0("run", run.num), bt_score
    )[
    , algorithm := sub("classif.", "", learner_id)]
  }
  batch_dt <- rbindlist(batch_dt_list)
  dcast(
    batch_dt, task_id + algorithm ~ run, value.var="classif.auc")
}
```

# Conclusion

We see that reproducibility is possible in mlr3.
Using `mlr3resampling::proj_grid()`

* Use tasks with `fold` column role.
* Default sets random seed for reproducible training.

Using `mlr3batchmark`, user needs to give `seed=1` argument to `batchtools::makeExperimentRegistry`.
