<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Simulations}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this vignette is to explain how to quantify the extent to
which it is possible to train on one data subset (say a geographic
region such as Europe), and predict on another data subset (say North
America). The ideas are similar to my previous blog posts about how to
do this in
[python](https://tdhock.github.io/blog/2022/generalization-to-new-subsets/) and
[R](https://tdhock.github.io/blog/2023/R-gen-new-subsets/).
Below we use data for simulated regression and classification problems.

## Simulated regression problems

Assume there is a data set with some rows from one person, some rows
from another,

```{r}
N <- 1000
library(data.table)
set.seed(1)
reg.dt <- data.table(
  x=runif(N, -2, 2),
  person=rep(1:2, each=0.5*N))
reg.pattern.list <- list(
  easy=function(x, person)x^2,
  impossible=function(x, person)(x^2+person*3)*(-1)^person)
reg.task.list <- list()
for(pattern in names(reg.pattern.list)){
  f <- reg.pattern.list[[pattern]]
  yname <- paste0("y_",pattern)
  reg.dt[, (yname) := f(x,person)+rnorm(N)][]
  task.dt <- reg.dt[, c("x","person",yname), with=FALSE]
  reg.task.list[[pattern]] <- mlr3::TaskRegr$new(
    pattern, task.dt, target=yname
  )$set_col_roles("person",c("group","stratum"))
}
reg.dt
```

The table above shows some simulated data for two regression problems:
* easy problem has the same pattern for each person, so it is possible/easy to train on one person, and accurately predict on another.
* impossible problem has a different pattern for each person, so it is impossible to train on one person, and accurately predict on another.

### Visualization

First we reshape the data using the code below,

```{r}
(reg.tall <- nc::capture_melt_single(
  reg.dt,
  pattern="easy|impossible",
  value.name="y"))
```

The table above is a more convenient form for the visualization which we create using the code below,

```{r}
if(require(ggplot2)){
  ggplot()+
    geom_point(aes(
      x, y),
      data=reg.tall)+
    facet_grid(
      pattern ~ person,
      labeller=label_both,
      space="free",
      scales="free")+
    scale_y_continuous(
      breaks=seq(-100, 100, by=2))
}
```

In the simulated data above, we can see that 
* for the easy pattern, it is the same for both people, so it should
  be possible/easy to train on one person, and accurately predict on
  another.
* for the impossible pattern, it is different for each person, so it
  should not be possible to train on one person, and accurately
  predict on another.
  
### Benchmark

In the code below, we define a 5-fold cross-validation experiment.

```{r}
(reg_same_other <- mlr3resampling::ResamplingSameOtherCV$new())
```

In the code below, we define two learners to compare,

```{r}
(reg.learner.list <- list(
  if(requireNamespace("rpart"))mlr3::LearnerRegrRpart$new(),
  mlr3::LearnerRegrFeatureless$new()))
```

In the code below, we define the benchmark grid, which is all
combinations of tasks (easy and impossible), learners (rpart and
featureless), and the one resampling method.

```{r}
(reg.bench.grid <- mlr3::benchmark_grid(
  reg.task.list,
  reg.learner.list,
  reg_same_other))
```

In the code below, we execute the benchmark experiment (in parallel
using the multisession future plan).

```{r}
if(require(future))plan("multisession")
(reg.bench.result <- mlr3::benchmark(reg.bench.grid))
```

The code below computes the test error for each split,

```{r}
reg.bench.score <- mlr3resampling::score(reg.bench.result)
reg.bench.score[1]
```

The code below visualizes the resulting test accuracy numbers.

```{r}
if(require(ggplot2)){
  ggplot()+
    scale_x_log10()+
    geom_point(aes(
      regr.mse, train.groups, color=algorithm),
      shape=1,
      data=reg.bench.score)+
    facet_grid(
      person ~ task_id,
      labeller=label_both,
      scales="free")
}
```

It is clear from the plot above that 
* for the easy task, training on same is just as good as all or other
  groups. rpart has much lower test error than featureless, in all
  three train groups.
* for the impossible task, the least test error is using rpart with same train groups; featureless with same train groups is next best; training on all is substantially worse (for both featureless and rpart); training on other is even worse (patterns in the two people are completely different).

## Simulated classification problems

Assume there is a data set with some rows from one person, some rows
from another,

```{r}
N <- 1000
library(data.table)
(full.dt <- data.table(
  label=factor(rep(c("spam","not spam"), l=N)),
  person=rep(1:2, each=0.5*N)
)[, signal := ifelse(label=="not spam", 0, 2)][])
```

Above each row has an person ID between 1 and 2. 
We can imagine a spam filtering system, that has training data for multiple people (here just two).
Each row in the table above represents a message which has been labeled as spam or not, by one of the two people.
Can we train on one person, and accurately predict on the other person?
To do that we will need some features, which we generate/simulate below:

```{r}
set.seed(1)
n.people <- length(unique(full.dt$person))
for(person.i in 1:n.people){
  use.signal.vec <- list(
    easy=rep(if(person.i==1)TRUE else FALSE, N),
    impossible=full.dt$person==person.i)
  for(feature.type in names(use.signal.vec)){
    use.signal <- use.signal.vec[[feature.type]]
    full.dt[
    , paste0("x",person.i,"_",feature.type) := ifelse(
      use.signal, signal, 0
    )+rnorm(N)][]
  }
}
full.dt
```

There are two sets of two features:

* For easy features, one is correlated with the label (`x1_easy`), and
  one is random noise (`x2_easy`), so the algorithm just needs to
  learn to ignore the noise feature, and concentrate on the signal
  feature. That should be possible given data from either person (same
  signal in each person).
* Each impossible feature is correlated with the label (when feature
  number same as person number), or is just noise (when person number
  different from feature number). So if the algorithm has access to
  the correct person (same as test, say person 2), then it needs to
  learn to use the corresponding feature `x2_impossible`. But if
  the algorithm does not have access to that person, then the best it
  can do is same as featureless (predict most frequent class label in
  train data).
  
### Visualization

Below we reshape the data to a table which is more suitable for visualization:

```{r}
(scatter.dt <- nc::capture_melt_multiple(
  full.dt,
  column="x[12]",
  "_",
  feature.type="easy|impossible"))
```

Below we visualize the pattern for each person and feature type:

```{r}
if(require(ggplot2)){
  ggplot()+
    geom_point(aes(
      x1, x2, color=label),
      shape=1,
      data=scatter.dt)+
    facet_grid(
      person ~ feature.type,
      labeller=label_both)
}
```

In the plot above, it is apparent that 

* for easy features (left), the two label classes differ in x1 values
  for both people. So it should be possible/easy to train on person 1, and
  predict accurately on person 2.
* for impossible features (right), the two people have different label
  patterns. For person 1, the two label classes differ in x1 values,
  whereas for person 2, the two label classes differ in x2 values. So
  it should be impossible to train on person 1, and predict accurately
  on person 2.

### Benchmark

We use the code below to create a list of classification tasks, for
use in the mlr3 framework.

```{r}
class.task.list <- list()
for(feature.type in c("easy","impossible")){
  feature.names <- grep(feature.type, names(full.dt), value=TRUE)
  task.col.names <- c(feature.names, "label", "person")
  task.dt <- full.dt[, task.col.names, with=FALSE]
  class.task.list[[feature.type]] <- mlr3::TaskClassif$new(
    feature.type, task.dt, target="label"
  )$set_col_roles(
    "person",c("group","stratum")
  )$set_col_roles(
    "label",c("target","stratum"))
}
class.task.list
```

Note in the code above that person is assigned roles group and
stratum, whereas label is assigned roles target and stratum.

The code below is used to define a 3-fold cross-validation experiment,

```{r}
(class_same_other <- mlr3resampling::ResamplingSameOtherCV$new())
```

The code below is used to define the learning algorithms to test,

```{r}
(class.learner.list <- list(
  if(requireNamespace("rpart"))mlr3::LearnerClassifRpart$new(),
  mlr3::LearnerClassifFeatureless$new()))
```

The code below defines the grid of tasks, learners, and resamplings.

```{r}
(class.bench.grid <- mlr3::benchmark_grid(
  class.task.list,
  class.learner.list,
  class_same_other))
```

The code below runs the benchmark experiment grid. Note that each
iteration can be parallelized by declaring a future plan.

```{r}
if(require(future))plan("multisession")
(class.bench.result <- mlr3::benchmark(class.bench.grid))
```

Below we compute scores (test error) for each resampling iteration,
and show the first row of the result.

```{r}
class.bench.score <- mlr3resampling::score(class.bench.result)
class.bench.score[1]
```

Finally we plot the test error values below.

```{r}
if(require(ggplot2)){
  ggplot()+
    geom_point(aes(
      classif.ce, train.groups, color=algorithm),
      shape=1,
      data=class.bench.score)+
    facet_grid(
      person ~ task_id,
      labeller=label_both,
      scales="free")
}
```

It is clear from the plot above that 
* for the easy task, training on same is just as good as all or other
  groups.
* for the impossible task, we must train on same group for minimal
  test error; training on all is almost as good, because the pattern
  in person 1 is orthogonal to person 2; training on other is just as
  bad as featureless, because patterns are different.

## Conclusion

In this vignette we have shown how to use mlr3resampling for comparing
test error of models trained on same/all/other groups.
