---
vignette: >
  %\VignetteIndexEntry{Using subset with group and stratum}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
litedown::reactor(
  collapse = TRUE,
  comment = "#>",
  fig.width=10,
  fig.height=6)
data.table::setDTthreads(1)
```

In `mlr3` a data set is represented as a `Task` with meta-data called roles which define how columns should be used for learning and evaluation.
The goal of this vignette is to explain how to use the new `subset` role with previous `group` and `stratum` roles.

# Introduction: role definitions

Previous column roles in `mlr3` are:

* `stratum`: proportions should be respected when splitting.
* `group`: rows that must stay together when splitting.

New column role proposed in `mlr3resampling` is:

* `subset`: one value is test, same/other/all values are train.

# Example using AZtrees data

The `AZtrees` data represent an image segmentation problem.
Each row is a pixel, which we want to classify as tree or not (`stratum` role).
Data were labeled by drawing polygons on a map (`group` role).
Data are divided into either 3 or 4 regions (`subset` role).

```{r}
data(AZtrees, package="mlr3resampling")
library(data.table)
options(datatable.print.keys=FALSE, datatable.print.class=FALSE)
AZtrees[, .(region3, polygon, y)]
```

Above we see the three columns that are relevant for these roles in these data.
Below we define a helper function which creates a `Task` and assigns roles based on arguments.

```{r}
trees <- function(..., cv=NULL, folds=3L){
  trees_task <- mlr3::as_task_classif(AZtrees, target="y")
  role_list <- list(...)
  for(role in names(role_list)){
    trees_task$col_roles[[role]] <- role_list[[role]]
  }
  if(is.null(cv))cv <- mlr3resampling::ResamplingSameOtherSizesCV$new()
  cv$param_set$values$folds <- folds
  cv$instantiate(trees_task)
  role_dt <- data.table(AZtrees)
  for(fold in 1:folds){
    set(role_dt, cv$test_set(fold), "fold", fold)
  }
  setkey(role_dt, polygon, fold)
  in_list <- list(
    rows=role_dt,
    groups=role_dt[, .(fold=paste(unique(fold), collapse=",")), by=.(y,polygon)])
  out_list <- list()
  for(unit in names(in_list)){
    in_dt <- in_list[[unit]]
    out_list[[unit]] <- rbind(in_dt[, table(y, fold)], TOTAL=in_dt[, table(fold)])
  }
  out_list
}
set.seed(1)
```

Above we set the seed to obtain reproducible results below (fold assignment is random).

## Cross-validation with no column roles

Below we show counts of rows and groups in each fold, when no column roles are set:

```{r}
trees()
```

We see in above that 

* in `rows`, there are about the same label distribution per fold (plus or minus 10 or 20 labels across folds), but not exactly the same (this can be achieved by setting `stratum` role). In `TOTAL` row we see the same number of rows per fold (plus or minus 1 because the total number of rows has a remainder when divided into the number of folds).
* in `groups`, rows from some polygons have been split into several different folds (`1,2` etc.), which is undesirable (and can be avoided by setting `group` role).

Below we do it again to emphasize the random variation that can occur (especially in label counts) when roles are not set:

```{r}
trees()
```

Below we compare with the previous `mlr3` code:

```{r}
trees(cv=mlr3::rsmp("cv"))
```

When no column roles are set, results are similar between previous `mlr3` and proposed `mlr3resampling`.

## Cross-validation on polygons

Below we compute counts when setting `group` role to `polygon`:

```{r}
trees(group="polygon")
```

Above we see 

* in `rows`, there is substantial label distribution variation across folds (count differences of about 100, which can be avoided by setting `stratum` role). We see the `TOTAL` number of rows is equal across folds (again plus or minus 1).
* in `group_counts` 
* in `groups`, each polygon has been assigned to one fold (no `1,2` etc.), which is desirable. We see different numbers of `TOTAL` groups per fold, which is expected, because the algorithm attempts to equalize number or rows per fold, and there are different numbers or rows per group.

Below we do it again to emphasize the random variation that can occur (especially in label counts) in this case:

```{r}
trees(group="polygon")
```

Above we see `TOTAL` is the same, but label counts per fold have changed.

Next, we try regular CV from mlr3,

```{r}
trees(group="polygon", cv=mlr3::rsmp("cv"))
```

We see a very different pattern: `TOTAL=63` polygons per group, but there is large variation in the number of rows per fold.

This comparison shows that the proposed `mlr3resampling` provides a new method for fold assignment using groups, that was not previously available in `mlr3`.


